### 2017-09-21 tex/full_2017/bibtemplate.tex
First commit of new paper-directory.

\documentclass[10pt]{IEEEtran}

\usepackage{times}
\usepackage[latin1]{inputenc}
\usepackage[hyphens]{url}
\usepackage[bookmarks={false},colorlinks={false},hyperref={false}]{hyperref}
\usepackage{nohyperref}

\begin{document}
\section{Introduction}
Examples for citations and references \cite{Firstman:2014:Approach} and \cite{Firstman:2013:Ongoing}
and \cite{Webpage:2014:Somewebpage}
and \cite{Bookman:2014:Somebook} and \cite{Chapman:2014:Starting}.
In a consecutive use these are referenced as
\cite{Firstman:2014:Approach,Firstman:2013:Ongoing,Webpage:2014:Somewebpage,Bookman:2014:Somebook,Chapman:2014:Starting}.

\bibliographystyle{IEEEtran}
\bibliography{bibtemplate_samples}

\end{document}